---
title: "Take-Home Final"
author: "Bartlomiej Morawski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
```{r}
load("ThreeCancers.RData")

# (i)
num.of.subj <- nrow(GeneExp) # number of subjects in the matrix
num.of.subj

num.of.genes <- ncol(GeneExp)
num.of.genes

# (ii)
count <- 0
a <- 1
for (i in CancerType){
  if (CancerType[a] == "LUAD"){
    count <- count + 1
    }
  a <- a + 1
}
count

# (iii)
mean(GeneExp[,2])

# (iv)
library(ggplot2)
ggplot(data=GeneExp, mapping=aes(x=GeneExp$Gene45, y=GeneExp$Gene68)) +
geom_point()

# (v)
cor.test(GeneExp$Gene45, GeneExp$Gene68)

```

## Problem 2

(i) Make a box plot of Gene15 vs CancerType and pass it the following arguemnts: col = c("red", "green", "blue") and ylab = "Gene15"
```{r}
plot(CancerType, GeneExp$Gene15, col = c("red", "green", "blue"), ylab = "Gene 15")
```

(ii) What does the red box represent?

The red box represents the correlation between the expression of gene 15 in the subjects and the subjects with BRCA cancer. 

(iii) Which cancer type(s) have outliers

Cancer's BRCA and KIRC have outliers.

(iv) Which cancer type has the smallest median expression level?

Cancer KIRC has the smallest median expression level. 

